---
title: "Study 1 - modelling: Reliability of the Serial Reaction Time task: If at first you don't succeed, try try try again"
author: "Catia Oliveira, Marianna E. Hayiou-Thomas, Lisa Henderson, University of York"
output:
    html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label = "allow")
```

```{r echo = T,warning=FALSE,message=FALSE}
library(rmarkdown)
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(influence.ME)
library(tidyverse)
```

### Functions: Holm-Bonferroni correction

```{r echo = F}

HB.correction <- function(alpha, p.value){
        n <- length(na.omit(p.value))
		   HB.pvalue <- alpha/(n - rank(p.value) + 1)
	       
}
```

# Data preprocessing

```{r echo = F}

# Read in file
Data <- read.csv("../data/Data_final_RT.csv", header = TRUE)
Data$logRT <- log(Data$RT)

#separate sessions

Session1_Data <- subset(Data, Session == 1)
Session2_Data <- subset(Data, Session == 2)

Session1_Prob <- subset(Session1_Data, Probability == 2)
Session1_Improb <- subset(Session1_Data, Probability == 1)
Session2_Prob <- subset(Session2_Data, Probability == 2)
Session2_Improb <- subset(Session2_Data, Probability == 1)

Data_results1 <- aggregate(RT ~Participant, Session1_Data, mean)
Data_results2 <- aggregate(RT ~Participant, Session2_Data, mean)

#------------------------------------

# PARTICIPANT OUTLIER REMOVAL BASED ON ACCURACY

#--- Same group

#Accuracy
pptmeanacc <-aggregate(Accuracy ~ Participant, data = Session1_Data, sum)
pptmeanacc2 <-aggregate(Accuracy ~ Participant, data = Session2_Data, sum)

#Z scores

pptmeanacc$AccZ <-scale(pptmeanacc$Acc, center = TRUE, scale = TRUE)
pptmeanacc2$AccZ <-scale(pptmeanacc2$Acc, center = TRUE, scale = TRUE)

# PARTICIPANT OUTLIER REMOVAL BASED ON Response Times
#---- Same group

# Mean response times 
pptmean_RT1 <-aggregate(RT ~ Participant, data= Session1_Data, mean)
pptmean_RT2 <-aggregate(RT ~ Participant, data= Session2_Data, mean)

####z scores
pptmean_RT1$RTZ <-scale(pptmean_RT1$RT, center = TRUE, scale = TRUE)
pptmean_RT2$RTZ <-scale(pptmean_RT2$RT, center = TRUE, scale = TRUE)

# plot response times
hist(Session1_Data$RT)
hist(Session2_Data$RT)

#Removal participants

Session1_trim1 <- subset(Session1_Data, Participant != "P083G01S02")
Session1_trim2 <- subset(Session1_trim1, Participant != "P028G02S02")
Session1_trim3 <- subset(Session1_trim2, Participant != "P076G02S01")
Session2_trim1 <- subset(Session2_Data, Participant != "P083G01S02")
Session2_trim2 <- subset(Session2_trim1, Participant != "P076G02S01")
Session2_trim3 <- subset(Session2_trim2, Participant != "P024G02S02")

#plot trimmed data

hist(Session1_trim1$RT)
hist(Session2_trim2$RT)

#rename datasets

Session1_finalData <- Session1_trim3
Session2_finalData <- Session2_trim3

# final dataset response times model
Data.trimmed <- rbind(Session1_finalData, Session2_finalData)
Data.trimmed2 <- subset(Data.trimmed, Epoch > 2)

```

# Descriptive statistics

```{r echo = F}
mean(Data.trimmed$RT)
sd(Data.trimmed$RT)

Desc <- Data.trimmed %>%
  group_by(Probability, Epoch, Session) %>%
  summarise(mean = mean(RT), sd = sd(RT))
```

# Contrasts

```{r echo = F}
#...........................Contrasts

#.......... Factor - Session
Data.trimmed$Session <- as.factor(Data.trimmed$Session)
contrasts(Data.trimmed$Session) <- c(1, -1)
contrasts(Data.trimmed$Session)

#.......... Factor - Epoch
Data.trimmed$Epoch <- as.factor(Data.trimmed$Epoch)
contrasts(Data.trimmed$Epoch) <- MASS::contr.sdif(5) 
contrasts(Data.trimmed$Epoch)

#.......... Factor - Probability
Data.trimmed$Probability <- as.factor(Data.trimmed$Probability)
contrasts(Data.trimmed$Probability) <- c(1, -1)
contrasts(Data.trimmed$Probability)
```

# Mixed effect models

## Max intercept model

```{r eval = F}
max.intercept.logmodel <- lmer(logRT ~ Epoch*Probability*Session + (1|Participant),data= Data.trimmed, REML=FALSE)
summary(max.intercept.logmodel)
```

### Checking assumptions

```{r eval = F}
#---------------Assumptions

par(mfrow=c(2,2))
#Absence of Homoskedasticity
plot(fitted(max.intercept.logmodel),residuals(max.intercept.logmodel))
abline(h = 0, col = "darkorange", lwd = 2)
#normality of residuals
hist(residuals(max.intercept.logmodel))
qqPlot(residuals(max.intercept.logmodel))

```

## Procedural learning models

### Model 1

```{r eval = F}
model1 <- lmer(logRT ~ Epoch*Probability*Session + (1 + Session * Epoch * Probability |Participant),data= Data.trimmed, REML=FALSE,
                    control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(model1)

```

### Model 2

```{r eval = F}
model2 <- lmer(logRT ~ Epoch*Probability*Session + (1 + Session * Epoch + Session * Probability + Epoch*Probability|Participant),data= Data.trimmed, REML=FALSE,
                    control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(model2)
```

### Model comparison

```{r eval = FALSE}

anova(model1, model2)
```

### Final model

```{r echo = F}
Model <- lmer(logRT ~ Epoch*Probability*Session + (1 + Session * Epoch + Probability |Participant),data= Data.trimmed, REML=FALSE,
              control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e+05))) #model selected 

res <- summary(Model)
print(summary(Model))

CI <- broom.mixed::tidy(Model, conf.int = TRUE)
#print(CI)

write.csv(CI, "RT1_results_with_influential.csv")

```

#### Influential cases

```{r echo = FALSE}

influentialcases.S1 <- influence(Model, "Participant")

Zdfbetas.S1 = as.data.frame(scale(dfbetas(influentialcases.S1)))

write.csv(Zdfbetas.S1, "zbetas-study1.csv")
```

#### Exclusion of influential participants

```{r echo = FALSE}

#Zdfbetas.S1 <- read.csv("zbetas-study1.csv")  %>% tibble::column_to_rownames(var = "X")

# Identify participants 
exclude.S1 <- Zdfbetas.S1 %>% 
  filter_all(any_vars(abs(.) > 3.29)) %>% 
  mutate(Participant = row.names(.))

exclude.S1 <- exclude.S1$Participant

write.csv(exclude.S1, "excluded_participants_study1.csv", row.names = FALSE)

# Exclude participants from dataset 
Data.final.S1 <- Data.trimmed[!Data.trimmed$Participant %in% exclude.S1,]

```

### Final model without influential cases

```{r echo = FALSE}

Model.S1 <- lmer(logRT ~ Epoch*Probability*Session + (1 + Session * Epoch + Probability |Participant),data= Data.final.S1, REML=FALSE,
              control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=1e+05))) #model with best fit

summary(Model.S1)

# Model parameters with CI
CI_S1_without_inf <- broom.mixed::tidy(Model.S1, conf.int = TRUE)
CI_S1_without_inf$HB.value <- HB.correction(.05, CI_S1_without_inf$p.value)
CI_S1_without_inf$sig <- ifelse(CI_S1_without_inf$p.value < CI_S1_without_inf$HB.value,"survive","does not survive")
write.csv(CI_S1_without_inf, "CI_S1_without_inf_march.csv")

```

## Models with similarity

```{r echo = F}
#----------------------------------------------------- Model 2 - similarity


# Contrast

#.......... Factor - Probability
Data.trimmed2$Probability <- as.factor(Data.trimmed2$Probability)
contrasts(Data.trimmed2$Probability) <- c(1, -1)
contrasts(Data.trimmed2$Probability)

#.......... Factor - Session
Data.trimmed2$Session <- as.factor(Data.trimmed2$Session)
contrasts(Data.trimmed2$Session) <- c(1, -1)
contrasts(Data.trimmed2$Session)


Data.trimmed2$logRT <- log(Data.trimmed2$RT)

#------------------------------------------------------Model building-------------------------------------
Data.trimmed2$Distance <- scale(Data.trimmed2$Distance, center = TRUE, scale = TRUE)

```

### Model similarity

```{r eval = F}
#Add probability and session as slope
Model_sim <- lmer(logRT ~ Session*Probability*Distance + (1 + Session * Probability |Participant),data= Data.trimmed2, REML=FALSE, control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(Model_sim)

```

### Model similarity 2

```{r eval = F}
#Add interaction between probability and session

Model_sim2 <- lmer(logRT ~ Probability*Session*Distance + (1 + Session + Probability |Participant),data= Data.trimmed2, REML=FALSE, control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(Model_sim2)
```

### Model comparison

```{r eval = F}
anova(Model_sim, Model_sim2) #Model_sim2
```

### Final model with similarity

```{r eval = F}
Model_sim_final <-  lmer(logRT ~ Probability*Session*Distance + (1 + Session * Probability |Participant),data= Data.trimmed2, REML=FALSE, control=lmerControl(optimizer="bobyqa")) #model with best fit

summary(Model_sim_final)
print(summary(Model_sim_final))

CI_sim <- broom.mixed::tidy(Model_sim_final, conf.int = TRUE)

print(CI_sim)
write.csv(CI_sim, "results_sim_with_influential1.csv")
```

#### Influential cases

```{r eval = FALSE}

influentialcases.sim <- influence(Model_sim_final, "Participant")

Zdfbetas.sim = as.data.frame(scale(dfbetas(influentialcases.sim)))

write.csv(Zdfbetas.sim, "zbetas-study1.sim.csv")
```

#### Exclusion of influential participants

```{r echo = FALSE}
#Zdfbetas.sim <- read.csv("zbetas-study1.sim.csv")  %>% tibble::column_to_rownames(var = "X")

# Identify participants 
exclude.S1.sim <- Zdfbetas.sim %>%
  filter_all(any_vars(abs(.) > 3.29)) %>% 
  mutate(Participant = row.names(.))

exclude.S1.sim <- exclude.S1.sim$Participant

write.csv(exclude.S1.sim, "excluded_participants_study1.sim.csv", row.names = FALSE)

# Exclude participants from dataset 
Data.final.sim <- Data.trimmed2[!Data.trimmed2$Participant %in% exclude.S1.sim,]

```

### Final model without influential cases

```{r echo = FALSE}

Model.S1.sim.final <-  lmer(logRT ~ Probability*Session*Distance + (1 + Session * Probability |Participant),data= Data.final.sim, REML=FALSE, control=lmerControl(optimizer="bobyqa")) #model with best fit

summary(Model.S1.sim.final)

# Model parameters with CI
CI_S1_without_inf.sim <- broom.mixed::tidy(Model.S1.sim.final, conf.int = TRUE)
CI_S1_without_inf.sim$HB.value <- HB.correction(.05, CI_S1_without_inf.sim$p.value)
CI_S1_without_inf.sim$sig <- ifelse(CI_S1_without_inf.sim$p.value < CI_S1_without_inf.sim$HB.value,"survive","does not survive")
write.csv(CI_S1_without_inf.sim, "CI_S1_without_inf.sim_march.csv")
```
