---
title: "# Supplementary experiment - Data analysis; Reliability of the Serial Reaction Time task: If at first you don't succeed, try try try again"
author: "Cátia Margarida Oliveira, Marianna E. Hayiou-Thomas, Lisa Henderson; University of York"
output:
    html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

### Importing libraries

```{r echo = T,warning=FALSE,message=FALSE}
library(tidyverse)
library(dplyr)
library(lme4)
library(ggplot2)
library(data.table)
library(lmerTest)
library(tidyr)
library(broom.mixed)
library(influence.ME)
```

# Holm-Bonferroni method

```{r}

HB.correction <- function(alpha, p.value){
       n <- length(na.omit(p.value))
		   HB.pvalue <- alpha/(n - rank(p.value) + 1)
	       
}
```

# SRT

```{r echo = F, results = 'hide',warning=FALSE,message=FALSE}

Data <- read.csv("../data/Data_SuppExp.csv", header = TRUE)

desc <- Data %>%
  group_by(Participant) %>%
  summarise(age = mean(Age), N = n())

#add missing columns-------

#assign 0 or 1 to odd or even trials
Data$Even_Odd <- rep(c("0", "1"), length.out=nrow(Data)) #even-odd column for split half

Data$RT <- Data$RT*1000 #multiply RT by 1000 for RT in milliseconds
Data$logRT <- log(Data$RT) #add log RT for modelling
Data$Group <- substr(as.character(Data$task), #add group column 
                     start= 1, 
                     stop= nchar(as.character(Data$task))-2)

#To estimate interval between sessions
Data <- tidyr::separate(Data, col = date, into = c("day","time"), sep = "_")

Interval <- Data %>% group_by(Participant, Session) %>% summarise(date = day[2])

Interval_wide <- reshape2::dcast(Interval, Participant ~ Session, 
                  value.var = c("date"))
names(Interval_wide)[names(Interval_wide) == "1"] <- "date1"
names(Interval_wide)[names(Interval_wide) == "2"] <- "date2"

Interval_wide$interval<- as.numeric(as.Date(Interval_wide$date2) - as.Date(Interval_wide$date1))
Interval_wide$date1 <- NULL
Interval_wide$date2 <- NULL
Data <- as.data.frame(list(Data, Interval_wide) %>%
                        reduce(full_join, by = "Participant"))

Data <- subset(Data, interval > 7)
#Data$AM_PM <- if_else(as.numeric(substr(Data$time,1,2))>5 & as.numeric(substr(Data$time,1,2)) < 16, "AM", "PM")


#divide data in separate groups for outlier detection -------------

ISI <- subset(Data, Group == "ISI")

Session1_Data <- subset(Data, Session == 1)
Session2_Data <- subset(Data, Session == 2)

nrow(Session1_Data)
nrow(Session2_Data)

Session1_Prob <- subset(Session1_Data, Probability == 2)
Session1_Improb <- subset(Session1_Data, Probability == 1)
Session2_Prob <- subset(Session2_Data, Probability == 2)
Session2_Improb <- subset(Session2_Data, Probability == 1)

#Outliers by group------------------------------------

# PARTICIPANT OUTLIER REMOVAL BASED ON ACCURACY

#ISI group-------------

ISI_1 <- subset(Session1_Data, Group == "ISI")
ISI_2 <- subset(Session2_Data, Group == "ISI")

#Accuracy
pptmeanacc_ISI <-aggregate(Accuracy ~ Participant, ISI_1, sum)
pptmeanacc2_ISI <-aggregate(Accuracy ~ Participant, ISI_2, sum)

#Z scores

pptmeanacc_ISI$AccZ <-scale(pptmeanacc_ISI$Accuracy, center = TRUE, scale = TRUE) #One outlier - BC48907
pptmeanacc2_ISI$AccZ2 <-scale(pptmeanacc2_ISI$Accuracy, center = TRUE, scale = TRUE) #One outlier - BC48907

#noISI Group -------------------

noISI_1 <- subset(Session1_Data, Group == "noISI")
noISI_2 <- subset(Session2_Data, Group == "noISI")

#Accuracy
pptmeanacc_noISI <-aggregate(Accuracy ~ Participant, noISI_1, sum)
pptmeanacc2_noISI <-aggregate(Accuracy ~ Participant, noISI_2, sum)

pptmeanacc_noISI$AccZ <- scale(pptmeanacc_noISI$Accuracy, center = TRUE, scale = TRUE) #two outliers - LWM92483, SF5990
pptmeanacc2_noISI$AccZ2 <- scale(pptmeanacc2_noISI$Accuracy, center = TRUE, scale = TRUE) #two outliers - LWM92483	and DO12606

#Removal outlier participants based on accuracy
#session1
ISI_1 <- subset(ISI_1, Participant != "BC48907")
noISI_1 <- subset(noISI_1, Participant != "LWM92483")
noISI_1 <- subset(noISI_1, Participant != "SF5990")

#session2
ISI_2 <- subset(ISI_2, Participant != "BC48907")
noISI_2 <- subset(noISI_2, Participant != "LWM92483")
noISI_2 <- subset(noISI_2, Participant != "DO12606")


# Participant outlier removal based on Response Times------------------------
#---- ISI

# Mean response times 
pptmean_RT1_ISI <-aggregate(RT ~ Participant, data= ISI_1, mean)
pptmean_RT2_ISI <-aggregate(RT ~ Participant, data= ISI_2, mean)

####z scores
pptmean_RT1_ISI$RTZ <-scale(pptmean_RT1_ISI$RT, center = TRUE, scale = TRUE) #one outlier - MC86247
pptmean_RT2_ISI$RTZ <-scale(pptmean_RT2_ISI$RT, center = TRUE, scale = TRUE) #one outlier - MC86247

#---- no ISI

# Mean response times 
pptmean_RT1_noISI <-aggregate(RT ~ Participant, data= noISI_1, mean)
pptmean_RT2_noISI <-aggregate(RT ~ Participant, data= noISI_2, mean)

####z scores
pptmean_RT1_noISI$RTZ <-scale(pptmean_RT1_noISI$RT, center = TRUE, scale = TRUE) #one outlier - EM77607 and MM57769
pptmean_RT2_noISI$RTZ <-scale(pptmean_RT2_noISI$RT, center = TRUE, scale = TRUE) #one outlier -EM77607

#Removal outlier participants
#session1
ISI_1 <- subset(ISI_1, Participant != "MC86247")
noISI_1 <- subset(noISI_1, Participant != "EM77607")

#session2
ISI_2 <- subset(ISI_2, Participant != "MC86247")
noISI_2 <- subset(noISI_2, Participant != "EM77607")

#remove participants who completed wrong task
Data.trimmed <- 


Data.trimmed <- rbind(ISI_1, ISI_2, noISI_1, noISI_2)

Data.trimmed2 <- subset(Data.trimmed, Epoch > 2)

```

# Models

## Set contrasts

```{r echo = F}

#Contrasts -----------------------------------------------------------------

#.......... Factor - Session
Data.trimmed$Session <- as.factor(Data.trimmed$Session)
contrasts(Data.trimmed$Session) <- cbind(c(1,-1))
contrasts(Data.trimmed$Session)

#.......... Factor - Epoch
Data.trimmed$Epoch <- as.factor(Data.trimmed$Epoch)
contrasts(Data.trimmed$Epoch) <- MASS::contr.sdif(5) 
contrasts(Data.trimmed$Epoch)

#.......... Factor - Probability
Data.trimmed$Probability <- as.factor(Data.trimmed$Probability)
contrasts(Data.trimmed$Probability) <- c(1, -1)
contrasts(Data.trimmed$Probability)

#.......... Factor - Group
Data.trimmed$Group <- as.factor(Data.trimmed$Group)
contrasts(Data.trimmed$Group) <- c(1, -1)
contrasts(Data.trimmed$Group)

Data.trimmed$Participant <- as.factor(Data.trimmed$Participant)
```


## RT.model

```{r eval=FALSE}
RT.model <- lmer(logRT ~ Probability * Epoch * Session * Group + (1+ Epoch * Session * Probability| Participant), data= Data.trimmed, REML = FALSE, control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(RT.model)
```

### Testing model fit

```{r eval = FALSE}

#Linearity

plot(residuals(RT.model), Data.trimmed$RT)

#Check for homoskedasticity

plot(fitted(RT.model),residuals(RT.model)) #shows homoskedasticity

# Check for normality of residuals
hist(residuals(RT.model), seq(min(residuals(RT.model)),max(residuals(RT.model)),by=((max(residuals(RT.model)) - min(residuals(RT.model)))/(length(residuals(RT.model))-1)))) #*data is skewed to the right

shapiro.test(residuals(RT.model)[1:5000])

#normality of residuals
hist(residuals(RT.model))
qqPlot(residuals(RT.model))

#DHARMa plotting
simulationOutput <- simulateResiduals(fittedModel = RT.model)
plot(simulationOutput)

plot(RT.model,sqrt(abs(residuals(.))) ~ fitted(.),type=c("p","smooth"))

```

## RT.model 2

```{r echo=FALSE}
RT.model2 <- lmer(logRT ~ Probability * Epoch * Session * Group + (1 + Epoch * Session + Probability * Session | Participant), data= Data.trimmed, REML = FALSE, control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(RT.model2)

CI.with.inf_S3 <- broom.mixed::tidy(RT.model2, conf.int = TRUE)
write.csv(CI.with.inf_S3, "../results/CI.with.inf_SuppExp.csv")

```

## Compare models

```{r eval=FALSE}

anova(RT.model, RT.model2)
```

### Influential cases

```{r echo=FALSE}
influentialcases <- influence(RT.model2, "Participant")
pchange(influentialcases)

#betas 
dfbetas(influentialcases)
plot(influentialcases, which="dfbetas")
Zdfbetas = as.data.frame(scale(dfbetas(influentialcases)))

write.csv(Zdfbetas, "../results/zbetas-SuppExp.csv")
```


### Exclusion of influential participants

```{r echo = FALSE}

#Zdfbetas <- read.csv("../results/zbetas-SuppExp.csv")  %>% tibble::column_to_rownames(var = "X")

exclude.S3 <- Zdfbetas %>% 
  filter_all(any_vars(abs(.) > 3.29)) %>% 
  mutate(Participant = row.names(.))

exclude.S3 <- exclude.S3$Participant

write.csv(exclude.S3, "../results/excluded_participants_SuppExp.csv", row.names = FALSE)

# Exclude participants from dataset 
Data.final3 <- Data.trimmed[!Data.trimmed$Participant %in% exclude.S3,]
```

## Final model

```{r echo = FALSE}

RT.model.final <- lmer(logRT ~ Probability * Epoch * Session * Group + (1+Epoch*Session+Probability*Session| Participant), data= Data.final3, REML = FALSE, control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

summary(RT.model.final)

# Model parameters with CI
CI_S3 <- broom.mixed::tidy(RT.model.final, conf.int = TRUE)

# Holm-Bonferroni correction
CI_S3$HB.value <- HB.correction(.05, CI_S3$p.value)
CI_S3$sig <- ifelse(CI_S3$p.value < CI_S3$HB.value,"survive","does not survive")
write.csv(CI_S3, "../results/CI_S3_without_inf.csv")

```

### Plotting fixed effects

```{r}

sjPlot::plot_model(RT.model.final, type = "int", mdrt.values = "meansd")

plot_model(RT.model.final, type = "pred", terms = c("Probability", "Epoch", "Group"))

```

## Age

## Contrasts

```{r echo = F}

#Age model --------------------------------------

#Contrasts

#.......... Factor - Session
Data.trimmed2$Session <- as.factor(Data.trimmed2$Session)
contrasts(Data.trimmed2$Session) <- cbind(c(1,-1))
contrasts(Data.trimmed2$Session)

#.......... Factor - Probability
Data.trimmed2$Probability <- as.factor(Data.trimmed2$Probability)
contrasts(Data.trimmed2$Probability) <- c(1, -1)
contrasts(Data.trimmed2$Probability)

#.......... Factor - Group
Data.trimmed2$Group <- as.factor(Data.trimmed2$Group)
contrasts(Data.trimmed2$Group) <- c(1, -1)
contrasts(Data.trimmed2$Group)

Data.trimmed2$Participant <- as.factor(Data.trimmed2$Participant)

Data.trimmed2$Age <- scale(Data.trimmed2$Age, center = TRUE, scale = TRUE)

```

## Age - RT model

```{r echo = FALSE}

Age.Model <- lmer(logRT ~ Probability*Session*Group*Age + (1 |Participant), data= Data.trimmed2, REML = FALSE)

```

## Age - RT model2

```{r echo = FALSE}

Age.Model2 <- lmer(logRT ~ Probability*Session*Group*Age + (1 + Probability*Session |Participant), data= Data.trimmed2, REML = FALSE, control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

summary(Age.Model2)

# Model parameters with CI
CI_S3.age <- broom.mixed::tidy(Age.Model2, conf.int = TRUE)
write.csv(CI_S3.age, "CI_SuppExp.age.csv")
```

## Age - RT model3

```{r echo = FALSE}

Age.Model3 <- lmer(logRT ~ Probability*Session*Group*Age + (1 + Probability + Session|Participant), data= Data.trimmed2, REML = FALSE, control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
summary(Age.Model3)

```

# Age Model comparison

```{r echo = FALSE}

anova(Age.Model, Age.Model2, Age.Model3)

```

## Influential cases

```{r echo = FALSE}

influentialcases.age <- influence(Age.Model2, "Participant")

Zdfbetas.age <- as.data.frame(scale(dfbetas(influentialcases.age)))

write.csv(Zdfbetas.age, "../results/zbetas-SuppExp.age.csv")
```

# Exclusion of influential participants

```{r echo = FALSE}

#Zdfbetas.age <- read.csv("../results/zbetas-SuppExp.age.csv")  %>% tibble::column_to_rownames(var = "X")

# Identify participants 
exclude.S3.age <- Zdfbetas.age %>% 
  filter_all(any_vars(abs(.) > 3.29)) %>% 
  mutate(Participant = row.names(.))

exclude.S3.age <- exclude.S3.age$Participant

write.csv(exclude.S3.age, "../results/excluded_participants_SuppExp.age.csv", row.names = FALSE)

# Exclude participants from dataset 
Data.final.age <- Data.trimmed2[!Data.trimmed2$Participant %in% exclude.S3.age,]

```

# Final model

```{r echo = FALSE}

age.model.final <- lmer(logRT ~ Probability*Session*Group*Age + (1 + Probability*Session |Participant), data= Data.final.age, REML = FALSE, control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

summary(age.model.final)

# Model parameters with CI
CI_S3_without_inf.age <- broom.mixed::tidy(age.model.final, conf.int = TRUE)

# Holm-Bonferroni correction
CI_S3_without_inf.age$HB.value <- HB.correction(.05, CI_S3_without_inf.age$p.value)
CI_S3_without_inf.age$sig <- ifelse(CI_S3_without_inf.age$p.value < CI_S3_without_inf.age$HB.value,"survive","does not survive")
write.csv(CI_S3_without_inf.age, "../results/CI_SuppExp_without_inf.age.csv")
```

